---
title: "My Report"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
source("compare_chart.R")
source("trend_chart.R")
source("map.R")
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

I chose to focus on the 'us-prison-pop.csv' for this assignment. Within
this, I wanted to focus on Washington State because that's where I am
from! I then weeded that down to only the most populated counties in
Washington State. The variables that I chose to work with included
female prison population, black population and black prison population.
I chose these variables as I believed I would find more relevant data
within marginalized groups.

### Summary Information

After sorting through data, I chose to find some averages to set the
tone for the research. Again, I focused on Washington State, and thus
chose to find averages among counties. The average population among
counties in 2016 was 426,861 people and the average prison population
was 994. Meaning, about 0.233% of all people in those counties were
incarcerated. As for women between 15-64 only, the average population
among Washington counties was 142,068. Their average population in
prison was 76, meaning 0.0535% women in those counties were
incarcerated. With 142,068/426,861 = 33.282% of the average population
but 76/994 = 7.646% of the average prison population, I am not
surprised. Anyone giving a quick look over the data could see that men
are incarcerated at much higher rates. Although women are a marginalized
group, it doesn't seem to apply to prison. I next chose to find the
highest and lowest values for female and black prison populations. I
found that the highest black population in prison occurred in King
County in 2002 and was 1,845. Out of 73,480, this means that 2.511% of
were incarcerated that year, much higher than the 2016 average of
0.233%. The highest female prison population occurred in Pierce County
in 2009 and was 297. Out of 272,547, this means that 0.109% were
incarcerated that year, double the average in 2016 of 0.0535%. The
lowest for both variables occurred simultaneously in Adams County in
1993 and was 0 for both. This year, Adams County only contained 14 total
inmates, thus it was an odd year. There must have been an event that
occurred that year that caused such a low prison population.

Average Total Population: 426860.5 Average Total Prison Population:
994.3636 Average Female Prison Population: 75.81818 Average Female
Population (15-64): 142068.1

Highest Black Prison Pop: 1845, King County 2002 Highest Female Prison
Pop: 297, Pierce County 2009 Lowest Black Prison Pop: 0, Adams County
1993 (14 total in prison) Lowest Female Prison Pop: 0, Adams County 1993
(14 total in prison)

### The Dataset

-Vera Institute collected the data -Vera uses the Incarceration Trends
data tool which uses data from the Bureau of Justice Statistics, Census
of Jails and Annual Survey of Jails. -The data was collected to confront
discrimination and disparities in the prison system as well as promote
the growth of it. -There are 153,811 rows ('us-prison-pop.csv'). -There
are 36 ('us-prison-pop.csv'). -You must ask yourself if there are any
disparities in the groups in and not in the prison population, as well
as who has the power to create these situations and actually place
citizens in prison? Lastly, how much missing data is there from
marginalized groups that were failed to be reported?

What are possible limitations or problems with this data? (at least 200
words)

-Immediately after loading the data into R, it is evident that there is
a lot of missing values replaced by NA that need to be filtered out
before finding any trends or coming to any conclusions. This was the
first thing I did before proceeding to embark on this assignment. Also
the data that I worked with in 'us-prison-pop.csv' was only updated to
2016, meaning there is 7 years of missing data. This last 7 years has
been monumental in the push and pull of racial tension and the pursuit
of equity. I believe that there could be an upward trend in the growth
of the prison system since 2016. Another limitation could be how the
data was collected. It was compiled from 3 official sources, but still
mass data like this can be very difficult to collect and organize
accurately. I would wager that there are definite errors in the numbers
I worked with, however this is the case for just about any data set one
may work with. I also noticed that when finding total populations for
each group of people, they only accounted for the ages between 15 and
64. Under 15 may be mostly irrelevant, but there are definitely
prisoners and members of the population who are liable to go to jail,
all over 64.

### Trends Over Time Chart

```{r, echo=FALSE}
female_prison_graph
```

For my trends over time chart, I included female prison populations and
how they developed over time in the largest four counties in Washington
State. I wanted to find if the marginalization of women had any hand in
the prison system. At no time did the number of inmates pass 300 and the
amounts seemed to sporadically jump up and down for all four locations:
King, Spokane, Pierce, and Snohomish. I ended up finding out that women
are incarcerated far less than men and there did not seem to be any
trend that suggested unfair imprisonment.


### Variable Comparison Chart

```{r, echo=FALSE}
king_county_line_graph
```

For my variable comparison chart I chose to evaluate the relationship between the black population of King County and its black population in prison. I wanted to find out if and slightly expected there to be a definite relation between the two and maybe even a steeper slope on the line representing the prisoners. However, I found that as the black population rapidly grew from 1990 to 2016, the population in prison stayed very steady. It seems that although the two variables are very much related, the population in prison does not seem to be affected much by the total population.

### Map

```{r, echo=FALSE}
map
```

For my map, I chose to represent the black prison population geographically. I used a color scale to show the density of black inmates for each state. I found that incarceration rates rise in coastal states and fall around the middle of the United States. I also found the most inmates to be in the Illinois/Detroit area. I was expecting more variation, however there is not much as the rates for incarceration are fairly similar among states, especially coastal states.

